### 2016-09-24 agda.tex
working on #6

% !TEX root = hazelnut-popl17.tex
So far, we have given an overview of the most important judgements and
rules in the semantics of Hazelnut, and stated the critical metatheorem,
Sensibility, and several auxiliary ``checks''. In a few cases, we have
informally sketched out why these metatheorems will hold. In order to
formally establish that our design meets our stated objectives, we have
mechanized the semantics of Hazelnut using the Agda proof
assistant \cite{norell:thesis} (also see the Agda Wiki, hosted
at \url{http://wiki.portal.chalmers.se/agda/}.)

This development is available in the supplemental material. At the time of
submission, the full set of rules is available and the proofs of
Theorem \ref{thrm:actsafe}, Theorem \ref{lemma:movement-erasure} and a
number of technical lemmas that we elided or mentioned informally in the
text are complete. We also have nearly all of the cases of Reachability and
Constructability complete, and we see no major barrier to the remaining
cases -- we prioritized developing a clear exposition of our contributions
after proving representative cases and finding that they were
straightforward. The remaining cases will be completed in the repository
linked in the non-anonymous supplementary material.

The documentation includes a more detailed discussion of the technical
representation decisions that we made. The main idea is standard: we encode
each judgement as a dependent type. The rules defining the judgements
become the constructors of this type, and derivations are terms of these
type. This is a rich setting that allows proofs to take advantage of
pattern matching on the shape of derivations, closely matching standard
on-paper proofs. No proof automation was used, because the proof structure
itself is likely to be interesting to researchers who plan to build upon
our work.

We adopt Barendregt's convention for bound
variables \cite{urban}. Hazelnut's semantics does not need substitution, so
we do not need to adopt more sophisticated encodings
(e.g. \cite{lh09unibind,Pouillard11}.)

\subsection{Determinism}
A useful property would be \emph{action determinism}, i.e. that performing
an action produces a unique result. Formally, this would be established as
follows:
\begin{enumerate}[itemsep=0px,partopsep=0px,topsep=0px]
\item If $\performTyp{\ztau}{\alpha}{\ztau'}$ and
$\performTyp{\ztau}{\alpha}{\ztau''}$ then $\ztau'=\ztau''$.
\item If $\hsyn{\hGamma}{\removeSel{\zexp}}{\htau}$ and
  $\performSyn{\hGamma}{\zexp}{\htau}{\alpha}{\zexp'}{\htau'}$ and
  $\performSyn{\hGamma}{\zexp}{\htau}{\alpha}{\zexp''}{\htau''}$ then
  $\zexp' = \zexp''$ and $\htau' = \htau''$.
\item If $\hana{\hGamma}{\removeSel{\zexp}}{\htau}$ and
  $\performAna{\hGamma}{\zexp}{\htau}{\alpha}{\zexp'}$ and
  $\performAna{\hGamma}{\zexp}{\htau}{\alpha}{\zexp''}$ then $\zexp' =
  \zexp''$.
\end{enumerate}

This is not a theorem of the system as described in this paper. The reason
is quite subtle: the semantics of two construction actions,
$\aConstruct{\fasc}$ and $\aConstruct{\flam{x}}$, behave differently in the
analytic case than they do in the synthetic case -- the difference has to
do with what the generated ascription contains. Both cases can technically
``fire'' due to action subsumption. The mechanization includes a ``proof''
of determinism for all cases but these two. Resolving this ambiguity
formally requires some rather inelegant machinery to exclude these
particular base cases from action subsumption, so we are content to
admonish implementors who apply action subsumption when another analytic
rule would do.