### 2016-11-02 acks.tex
added an acknowledgement for Danny's COPE grant

Thanks to Ed Morehouse for counsel on mechanization and Barendrecht's
Convention, and to Vincent Zeng for \textit{pro bono} artistic services.
%This is Danny Dig's COPE grant. He asked me to add the grant.
This work was partially funded through the NSF CCF-1439957 grant.