### 2016-11-07 hazelnut-popl17.tex
revised abstract and intro

\def\OPTIONConf{1}%
\newif\iftr\trtrue
\RequirePackage{etex}
\documentclass[9pt]{sigplanconf}
\toappear{}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ stmaryrd }
\usepackage{mathpartir}
\usepackage{stmaryrd}
\usepackage{wasysym}
\usepackage{multicol}
\usepackage{extarrows}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\definecolor{light-gray}{gray}{0.9}
\usepackage{soul}
\setulcolor{red}
\usepackage{mathpazo}
\usepackage{colortab}
\usepackage[lowtilde]{url}
\usepackage{todonotes}
\usepackage{listings}
\lstset{tabsize=2,
basicstyle=\ttfamily\fontsize{8pt}{1em}\selectfont}
\usepackage{microtype}
\sloppy
\def \TirNameStyle #1{\small\rulename{#1}}
\renewcommand{\MathparLineskip}{\lineskiplimit=.3\baselineskip\lineskip=.3\baselineskip plus .2\baselineskip}

\usepackage{joshuadunfield}
\usepackage{llproof}
\usepackage{rulelinks}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{definition}{Definition}
\newenvironment{proof-sketch}{\noindent{\emph{Proof Sketch.}}}{\qed}
\makeatletter

\renewcommand\topfraction{0.85}
\renewcommand\bottomfraction{0.85}
\renewcommand\textfraction{0.1}
\renewcommand\floatpagefraction{0.85}

\AtBeginDocument{%
 \abovedisplayskip=2pt
 \abovedisplayshortskip=0pt
 \belowdisplayskip=2pt
 \belowdisplayshortskip=0pt
}

\setlength{\floatsep}{10pt}
\setlength{\textfloatsep}{12pt}

\usepackage[compact]{titlesec}
\titlespacing*{\section}{0pt}{4pt}{2pt}
\titlespacing*{\subsection}{0pt}{4pt}{2pt}
\titlespacing*{\subsubsection}{0pt}{4pt}{2pt}
\titlespacing*{\paragraph}{0pt}{4pt}{2pt}
\setlength{\skip\footins}{3pt plus 1px minus 5px}

% known bug: http://tex.stackexchange.com/questions/1522/pdfendlink-ended-up-in-different-nesting-level-than-pdfstartlink
\usepackage[hidelinks,breaklinks,draft=false]{hyperref}

\usepackage{enumitem}
\makeatletter
\def\thm@space@setup{%
  \thm@preskip=4px plus 2px minus 2px
  \thm@postskip=\thm@preskip % or whatever, if you don't want them to be equal
}
\makeatother

\usepackage{todonotes}
\usepackage{xcolor}

\input{macros}
\begin{document}

\conferenceinfo{-}{-}
\copyrightyear{-}
\copyrightdata{[to be supplied]}

\preprintfooter{Draft}   % 'preprint' option specified.

\title{Hazelnut: A Bidirectionally Typed \\ Structure Editor
 Calculus}
\iftr
\authorinfo{
        Cyrus Omar$^1$
        \and Ian Voysey$^1$
        \and Michael Hilton$^2$
        \and Jonathan Aldrich$^1$
        \and Matthew A. Hammer$^3$
}
{
 $^1$~Carnegie Mellon University (USA)
 \and
 $^2$~Oregon State University (USA)
 \and
 $^3$~University of Colorado Boulder (USA)\vspace{-10px}
}
{
%\email{comar@cs.cmu.edu}\\
%\texttt{http://www.cs.cmu.edu/\homedir comar/}
%\email{hiltonm@eecs.oregonstate.edu}\\
}
\else
\authorinfo{~}{~}{\vspace{-10px}}
\fi

\maketitle
\begin{abstract}

\emph{Structure editors} allow programmers to edit the tree structure of
a program directly. This can have cognitive benefits, particularly for
novice and end-user programmers. It also simplifies matters for tool
designers, because they do not need to contend with malformed program text.

This paper introduces Hazelnut, a {structure editor} based on a small
bidirectionally typed lambda calculus extended with \emph{holes} and
a \emph{cursor}. Hazelnut goes one step beyond
syntactic well-formedness: its {edit actions} operate over statically
meaningful incomplete terms.  Na\"ively, this would force the programmer to construct terms in a 
rigid ``outside-in'' manner. To avoid this problem, the {action semantics}
automatically places terms assigned a type that is inconsistent with the
expected type \emph{inside} a {hole}. This meaningfully defers the type consistency
check until the term inside the hole is \emph{finished}.

Hazelnut is not intended as an end-user tool itself. Instead, it serves as a foundational account of typed structure
editing. To that end, we describe how
Hazelnut's rich metatheory, which we have mechanized using the Agda proof assistant, serves as a guide when we extend the calculus to include binary sum types. We also discuss various interpretations of holes, and in so doing
reveal connections with gradual typing and contextual modal type theory
(the Curry-Howard interpretation of contextual modal logic.) Finally, we
discuss how Hazelnut's semantics lends itself to implementation as an event-based functional reactive program. Our simple reference implementation is written in OCaml and compiled using \lstinline{js_of_ocaml}.

\end{abstract}

\category{D.3.1}{Programming Languages}{Formal Definitions and Theory}
\category{D.2.3}{Software Engineering}{Coding Tools and Techniques}[Program Editors]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
\keywords
structure editors, bidirectional type systems, gradual typing, mechanized metatheory

\section{Introduction}\label{sec:introduction}
\input{intro}

\section{Programming in Hazelnut}\label{sec:example}
\input{example}

\section{Hazelnut, Formally}
\label{sec:hazel}
\input{formally}

\subsection{Mechanization}
\label{sec:mech}\label{sec:mt}
\input{agda}

\section{Extending Hazelnut}\label{sec:extending}
\input{sumtypes}

\section{Implementation}
\label{sec:impl}
\input{implementation}

\section{Related Work and Discussion}\label{sec:rw}
\todo{move into own file}
\subsection{Structure Editors}
Syntactic structure editors have a long history -- the Cornell Program
Synthesizer~\cite{teitelbaum_cornell_1981} was first introduced in
1981. Novice programmers have been a common target for structure
editors. For example, GNOME~\cite{garlan_gnome:_1984} was developed to
teach programming to undergraduates.
Alice~\cite{Conway:2000:ALL:332040.332481} is a 3-D programming language
with an integrated structure editor for teaching novice CS undergraduate
students. Scratch~\cite{Resnick:2009:SP:1592761.1592779} is a structure
editor targeted at children ages 8 to 16.
TouchDevelop \cite{tillmann_touchdevelop:_2011} incorporates a structure
editor for programming on touch-based devices, and is used to teach high
school students. An implementation of Hazelnut might be useful in teaching
students about the typed lambda calculus, though that has not been our
explicit aim with this work.

Not all structure editors are for educational purposes. For example,
mbeddr \cite{voelter_mbeddr:_2012} is a structure editor for a C-based
programming language (nominally, for programming embedded systems.)
Lamdu~\cite{lamdu}, like Hazelnut, is a structure editor for a statically
typed functional language. It is designed for use by professional
programmers.

The examples given so far either define a language with a trivial static
semantics, or do not attempt to maintain well-typedness as an edit
invariant. This can pose problems, for reasons discussed in the
Introduction. One apparent exception is Unison~\cite{unison}, a structure
editor for a typed functional language similar to Haskell. Like Hazelnut,
it seems to define some notion of well-typedness for expressions with holes
(though there is no technical documentation on virtually any aspect of its
design.) Unlike Hazelnut, it does not have a notion analogous to Hazelnut's
notion of a non-empty hole. As such, programmers must construct programs in
a rigid outside-in manner, as discussed in Sec. \ref{sec:example}. Another
system with the same problem is Polymorphic Blocks, a block-based user
interface where the structure of block connectors encodes a
type \cite{DBLP:conf/chi/LernerFG15}.

We fundamentally differ from these projects in our design philosophy: we
consider it essential to start by building type theoretic foundations,
which are independent of nearly all decisions about the user interface. In
contrast, these editors have developed innovative user interfaces (e.g. see
the discussion in \cite{DBLP:conf/sle/VolterSBK14}) but lack a principled
foundational calculus. In this respect, we follow the philosophical
approach taken by languages that are rooted in the type theoretic tradition
and have gone to great effort to develop a clear metatheory, like Standard
ML \cite{mthm97-for-dart,Harper00atype-theoretic}.  In the future, we hope
that these lines of research will merge to produce a human-usable typed
structure editor with sound formal foundations. Our contribution, then, is
in defining and analyzing the theoretical properties of a small
foundational calculus that could be extended to achieve this vision.

Some structure editor generators do operate on formal or semi-formal
definitions of an underlying language. For example, the Synthesizer
Generator~\cite{Reps:1984:SG:390010.808247} allows the user to define an
attribute grammar-based language implementation that then can be used to
generate a structured editor. CENTAUR~\cite{Borras:1988:CS:64140.65005}
produces a language specific environment from a user defined formal
specification of a language. Barista is a programmatic toolkit for building
structure editors \cite{ko_barista:_2006}. mbeddr is built on top of the
commercial JetBrains MPS framework for constructing structure
editors \cite{voelter2011language,DBLP:journals/software/VoelterWK15}. These
systems do not give a semantics to the edit states of the structure editor
itself, or maintain non-trivial edit invariants, as Hazelnut does.

Related to structure editors are value editors, which operate directly on
simple values (but not generally expressions or functions) of a programming
language. For example, Eros is a typed value editor based in
Haskell \cite{DBLP:conf/icfp/Elliott07}.

\subsection{Gradual Type Systems}
A significant contribution of this paper is the discovery of a clear
technical relationship between typed structure editing and gradual
typing. In particular, the machinery necessary to give a reasonable
semantics to type holes -- i.e. type consistency and type matching --
coincides with that developed in the study of gradual type systems for
functional languages. The pioneering work of Siek and Taha \cite{Siek06a}
introduced type consistency. Subsequent work developed the concept of type
matching \cite{DBLP:conf/popl/RastogiCH12,DBLP:conf/popl/GarciaC15}. In
retrospect, this relationship is perhaps unsurprising: gradual typing is,
notionally, also motivated by the idea of iterated development of a program
where every intermediate state is well-defined in some sense.

Recent work has discovered a systematic procedure for generating a
``gradual version'' of a standard type
system \cite{DBLP:conf/popl/CiminiS16}. This system, called the
Gradualizer, operates on a logic program that specifies a simple type
assignment system with some additional annotations to generate a
corresponding specification of a gradual type system. The authors leave
support for working with bidirectional type systems as future work. This
suggests the possibility of an analogous ``Editorializer'' that generates a
specification of a typed structure editor from a simple language
definition. Our exposition in Sec. \ref{sec:extending} certainly suggests
that many of the necessary definitions follow seemingly mechanically from
the definition of the static semantics, and the relationship with gradual
typing suggests that many of the technical details of this transformation
may already exist in the Gradualizer. One possibility we have explored
informally is to use Agda's reflection features to implement such a system.

An aspect of gradual typing that we did not touch on directly here is its
concern with assigning a dynamics to programs where type information is not
known, by inserting dynamic type casts. This would correspond to assigning
a dynamics to Hazelnut expressions with type holes such that a run-time
error occurs when a type hole is found to be unfillable through evaluation.

\subsection{Type Holes as Unification Variables}
Another possible interpretation of type holes is as explicit unification
variables. In other words, we might define the dynamics of Hazelnut such
that a program cannot be run if some automated type inference procedure
cannot statically fill in any type holes that arise. For a simple calculus,
e.g. the STLC upon which Hazelnut is based, type inference is known to be
decidable \cite{damas1982principal}. In more complex settings, e.g. in a
dependently typed language, a partial decision procedure may still be
useful in this regard, both at edit-time and (just prior to)
run-time. Indeed, text editor modes for proof assistants, e.g. for Agda,
attempt to do exactly this for indicated ``type holes'' (and do not always
succeed.)

\subsection{Exceptions}
Expression holes, too, could dynamically be interpreted in several
ways. One straightforward approach would be to dynamically raise an
exception if evaluation encounters an expression hole. Indeed,
placing \textt{raise Unimplemented} or similar in portions of an expression
that are under construction is a common practice across programming
languages.

\subsection{Type-Directed Program Synthesis}
Some text editor modes, e.g. those for proof assistants like
Agda \cite{norell:thesis} and Idris \cite{brady2013idris}, support a more
explicit hole-based programming model where indicated expression holes are
treated as sites where the system can be asked to automatically generate an
expression of an appropriate type. These systems are not statically
well-defined themselves (though see below.)

The Graphite system used an informal model of typed holes in Java to allow
library providers to associate interactive code generation interfaces with
types \cite{Omar:2012:ACC:2337223.2337324}.

More generally, the topic of type-directed program synthesis an active area
of research, e.g. \cite{DBLP:conf/pldi/OseraZ15}. By maintaining static
well-definedness throughout the editing process, Hazelnut provides
researchers interested in editor-integrated type-directed program synthesis
with a formal foundation upon which to build.

\subsection{Contextual Modal Type Theory}
Expression holes can also be understood by invoking the notion of
a \emph{metavariable} as found in contextual modal type theory
(CMTT) \cite{DBLP:journals/tocl/NanevskiPP08}. In particular, expression
holes have types and are surrounded by contexts, just as metavariables in
CMTT are associated with types and contexts. This helps to clarify the
logical meaning of a typing derivation in Hazelnut -- it conveys
well-typedness relative to an (implicit) modal context that extracts each
expression hole's type and context. The modal context must be emptied --
i.e. the expression holes must be instantiated with expressions of the
proper type in the proper context -- before the expression can be
considered complete. This corresponds to the notion of modal necessity in
contextual modal logic.

Making the modal context explicit in our semantics is not technically
useful given our stated purpose -- interactive program editing is not
merely hole filling in Hazelnut (i.e. the cursor need not be on a
hole). Moreover, the hole's type and context become apparent as our action
semantics traverses the zipper structure on each step. Some interactive
proof assistants, however, support a tactic model based directly on hole
filling, so the connection to CMTT and similar systems is more useful. For
example, Beluga \cite{DBLP:conf/flops/Pientka10} is based on dependent CMTT
and aspects of Idris' editor support \cite{brady2013idris} are based on
McBride's OLEG \cite{mcbride2000dependently} and Lee and Friedman have
explored a lambda calculus with contexts for a similar
purpose \cite{DBLP:conf/icfp/LeeF96}.

One interesting avenue of future work is to elaborate expression holes to
CMTT's closures, i.e. CMTT terms of the form
$\mathsf{clo}(u; \text{id}(\Gamma))$ where $u$ is a unique metavariable
associated with each hole and $\text{id}(\Gamma)$ is the explicit identity
substitution. This would allow us to evaluate expressions with holes such
that the closure ``accumulates'' substitutions explicitly. When evaluation
gets ``stuck'' (as it can, for CMTT does not define a dynamics equipped
with a notion of progress under a non-empty modal context), it would then
be possible for the programmer to choose holes from amongst the visible
holes (which may have been duplicated) to edit in their original
context. Once finished, the CMTT hole instantiation operation, together
with a metatheorem that establishes that reduction commutes with
instantiation, would enable an ``edit and resume'' feature with a clear
formal basis. This notion of reduction commuting with instantiation has
also been studied in other
calculi \cite{DBLP:journals/entcs/Sands97}. Being able to edit a running
program also has connections to less formal work on ``live programming''
interfaces \cite{burckhardt2013s,lamdu}.

\subsection{Evaluation Strategies: A High-Dimensional Space}
To summarize, we have discussed three different evaluation strategies in
the presence of type holes:
\begin{enumerate}[noitemsep]
\item ...as preventing evaluation (the standard approach.)
\item ...as unknown types, in the gradual sense.
\item ...as unification variables.
\end{enumerate}
In addition, we have discussed four different evaluation strategies in the
presence of expression holes:
\begin{enumerate}[noitemsep]
\item ...as preventing evaluation (the standard approach.)
\item ...as causing exceptions.
\item ...as sites for automatic program synthesis.
\item ...as the closures of CMTT.
\end{enumerate}

Every combination of these choices could well be considered in the design
of a full-scale programming system in the spirit of Hazelnut. Indeed, the
user might be given several options from among these combinations,
depending on their usage scenario. Many of these warrant further inquiry.

\section{Conclusion}
\label{sec:future}
This paper presented Hazelnut, a type theoretic structure editor
calculus. Our aim was to take a principled approach to its design by
formally defining its static semantics as well as its action semantics and
developing a rich metatheory. Moreover, we have mechanized substantial
portions of the metatheory, including the crucial Sensibility theorem that
establishes that every edit state is statically meaningful.

In addition to simplifying the job of an editor designer, typed structure
editors also promise to increase the speed of development by eliminating
redundant syntax and supporting higher-level primitive actions. However, we
did not discuss such ``edit costs'' here, because they depend on particular
implementation details, e.g. whether a keyboard or a mouse is in
use. Indeed, we consider it a virtue of this work that such implementation
details do not enter into our design.

\subsection{Future Work}
\subsubsection{Richer Languages}
Hazelnut is, obviously, a very limited language at its core. So the most
obvious avenue for future work is to increase the expressive power of this
language by extension. Our plan is to simultaneously maintain a
mechanization and implementation (following, for example, Standard ML) as
we proceed, ultimately producing the first large-scale, formally verified
bidirectionally typed structure editor.

It is interesting to note that the demarcation between the language and the
editor is fuzzy (indeed, non-existent) in Hazelnut. There may well be
interesting opportunities in language design when the language is being
codesigned with a typed structure editor. It may be that certain language
features are unnecessary given a sufficiently advanced type-aware structure
editor (e.g. SML's \texttt{open}?), while other features may only be
practical with editor support. We intend to use Hazelnut and derivative
systems thereof as a platform for rigorously exploring such questions.

\subsubsection{Editor Services}
There are various aspects of the editor model that we have not yet
formalized. For example, our action model does not consider how actions are
actually entered using, for example, key combinations or chords. In
practice, we would want also to rank available actions in some reasonable
manner (perhaps based on usage data gathered from other users or code
repositories.) Designing a rigorous typed probability model over actions
and H-expressions is one avenue of research that we have started to
explore, with intriguing initial results.

\subsubsection{Programmable Actions}
Our language of actions is intentionally primitive. However, even now it
acts much like a simple imperative command language. This suggests future
expansion to, for example, a true tactic language. Alternatively, it may be
more useful to develop the notion of an \emph{action macro}, whereby
functional programs could themselves be lifted to the level of actions to
compute non-trivial compound actions. Such compound actions would give a
uniform description of transformations ranging from the simple -- like
``move the cursor to the next hole to the right'' -- to quite complex whole
program refactoring, while remaining subject to the core Hazelnut
metatheory.

\subsubsection{Views}
Another research direction is in exploring how types can be used to control
the presentation of expressions in the editor. For example, following an
approach developed in a textual setting of developing \emph{type-specific
languages} (TSLs), it should be possible to have the type that an
expression is being analyzed against define alternative display forms and
interaction modes \cite{TSLs}.

It should also be possible to develop a semantics of semantic comments,
i.e. comments that mention semantic structures. These would be subject to
the same operations, e.g. renaming, as other structures, helping to address
the problem of comments becoming inconsistent with code.

\subsubsection{Collaborative Programming}
Finally, we did not consider any aspects of \emph{collaborative
programming}, such as a packaging system, a differencing algorithm for use
in a source control system, support for multiple simultaneous focii for
different users, and so on. These are all interesting avenues for future
work.

\subsubsection{Empirical Evaluation}
Although we make few empirical claims in this paper, it is ultimately an
empirical question as to whether structure editors, and typed structure
editors, are practical. We hope to conduct user studies once a richer
semantics has been developed.

\subsubsection{More Theory}
Connections with gradual type systems and CMTT, discussed in the previous
section, seem likely to continue to be revealing.

The notion of having one of many possible locations within a term under a
cursor has a very strong intuitive connection to the proof theoretic notion
of focusing \cite{Simmons11tr}. Building closer connections with proof
theory (and category theory) is likely to be a fruitful avenue of further
inquiry.

\begin{quote}
\emph{In any case, these are but steps toward more graphical program-description
systems, for we will not forever stay confined to mere strings of symbols.}

--- Marvin Minsky, Turing Award lecture \cite{DBLP:journals/jacm/Minsky70}
\end{quote}

\clearpage

\bibliographystyle{abbrvnat}
% The bibliography should be embedded for final submission.
\bibliography{bibliography}

\iftr
\clearpage
%\onecolumn
\appendix
\input{appendix}
\else
% No Appendix Here
\fi

\end{document}